---
title: "Homework 2"
author: "Ben Rodwell"
date: "February 23, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Every Saturday, at the same time, a primatologist goes and sits in the forest in the morning and listens for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number calls she will hear in that time is 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.

## 1.a. What is the probability that she will hear more than 8 calls during any given session?

```{r}
ppois(q = 8, lambda = 15, lower.tail = FALSE)  
```

## 1.b. What is the probability that she will hear no calls in a session?
```{r}
ppois(q = 0, lambda = 15, lower.tail = T)
dpois(x = 0, lambda = 15)  
```

## 1.c. What is the probability that she will hear exactly 3 calls in a session?
```{r}
dpois(x = 3, lambda = 15)
```

## 1.d. Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.
```{r}
x <- 0:30
l = 15
poistiti <- dpois(x = 0:30, lambda = 15)
barplot(poistiti, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X = x)",
        main = paste0("Probability Mass Function of Titi Monkey calls\nlambda = ", l), xlim = c(0,30))
```

## 1.e. Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions)
```{r}
rpois(n = 104, lambda = 15)
```

## 1.f. Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?
```{r}
randopois <- rpois(n = 104, lambda = 15)
hist(x = randopois, xlim = c(0,30))
```

