---
title: "Homework 2"
author: "Ben Rodwell"
date: "February 23, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# QUESTION 1: Every Saturday, at the same time, a primatologist goes and sits in the forest in the morning and listens for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number calls she will hear in that time is 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.

## 1.a. What is the probability that she will hear more than 8 calls during any given session?

```{r}
ppois(q = 8, lambda = 15, lower.tail = FALSE)  
```

## 1.b. What is the probability that she will hear no calls in a session?
```{r}
ppois(q = 0, lambda = 15, lower.tail = T)
dpois(x = 0, lambda = 15)  
```

## 1.c. What is the probability that she will hear exactly 3 calls in a session?
```{r}
dpois(x = 3, lambda = 15)
```

## 1.d. Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.
```{r}
x <- 0:30
l = 15
poistiti <- dpois(x = 0:30, lambda = 15)
barplot(poistiti, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X = x)",
        main = paste0("Probability Mass Function of Titi Monkey calls\nlambda = ", l), xlim = c(0,30))
```

## 1.e. Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions)
```{r}
rpois(n = 104, lambda = 15)
```

## 1.f. Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?
```{r}
randopois <- rpois(n = 104, lambda = 15)
hist(x = randopois, xlim = c(0,30))
```



# QUESTION 2: Load in the dataset “zombies.csv” from my GitHub repository at https://github.com/difiore/ADA-2019. This data includes the first and last name and gender of the entire population of 1000 people who have survived the zombie apocalypse and are now ekeing out an existence somewhere on the East Coast, along with several other variables (height, weight, age, number of years of education, number of zombies they have killed, and college major see here for info on important post-zombie apocalypse majors.

## Load in the zombies.csv dataset
```{r}
library(tidyverse)
library(curl)
library(manipulate)
library(ggplot2)
library(dplyr)
f <- f <- curl("https://raw.githubusercontent.com/difiore/ADA-2019/master/zombies.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```


# 2.a. Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education).

## A function to calculate population variance
```{r}
popvar <- function(x) {
  sum((x - mean(x))^2)/(length(x))}
popvar(x)
```

## A function to calculate polulation standard deviation
```{r}
popsd <- function(x) {
  sqrt(popvar(x))}
popsd(x)
```

## Calculate Mean for each variable
```{r}
mean(d$height)
mean(d$weight)
mean(d$age)
mean(d$zombies_killed)
mean(d$years_of_education)
```


## calculate SD for each variable
```{r}
popsd(d$height)
popsd(d$weight)
popsd(d$age)
popsd(d$zombies_killed)
popsd(d$years_of_education)
```











# 2.b. Use {ggplot} and make boxplots of each of these variable by gender.
```{r}
library(ggplot2)
heightbox <- ggplot(d, aes(x = gender, y = height)) + 
  geom_boxplot()
heightbox

weightbox <- ggplot(d, aes(x = gender, y = weight)) + 
  geom_boxplot()
weightbox

agebox <- ggplot(d, aes( x = gender, y = age)) + 
  geom_boxplot()
agebox

zombieskilledbox <- ggplot(d, aes(x = gender, y = zombies_killed)) + 
  geom_boxplot()
zombieskilledbox

yearsofeducationbox <- ggplot( d, aes(x = gender, y = years_of_education)) +
  geom_boxplot()
yearsofeducationbox
```









# 2.c. Use {ggplot} and make scatterplots of height and weight in relation to age. Do these variables seem to be related? In what way?
```{r}
heightfit = lm(d$height~d$age)
summary(heightfit)
heightscatter <- ggplot(d, aes(x= age, y= height)) + geom_point() + 
  geom_abline(mapping = NULL, data = NULL, slope =  0.94251,
  intercept = 48.73566, size = 2, col= "forestgreen") + labs(title = "Height versus age of survivors")
heightscatter

```

## Height and age have somewhat a slightly negatively allometric relationship. Height increases at a relatively slower rate than age does.

```{r}


weightfit = lm(d$weight~d$age)
summary(weightfit)
weightscatter <- ggplot(d, aes(x= age, y= weight)) + geom_point() + 
  geom_abline(mapping = NULL, data = NULL, slope =  1.9879,
  intercept = 104.0563, size = 2, col= "purple") + labs(title = "Weight versus age of survivors")
weightscatter

```

## Weight and age have strongly positive allometric relationship. Weight increases at a relatively quicker rate than age does.


# 2.d. Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not? For those that are not, can you determine what common distribution they are drawn from?

## Height
```{r}
qqnorm(d$height, main = "Normal QQ plot Height")
qqline(d$height, col = "gray")
```

### Height appears to have a normal distribution

## Weight
```{r}
qqnorm(d$weight, main = "Normal QQ plot Weight")
qqline(d$weight, col = "gray")
```

### Weight appears to have a normal distribution

## Age
```{r}
qqnorm(d$age, main = "Normal QQ plot Age")
qqline(d$age, col = "gray")
```

### Age appears to have a normal distribution

## Zombies killed
```{r}
qqnorm(d$zombies_killed, main = "Normal QQ plot # of Zombies Killed")
qqline(d$zombies_killed, col = "gray")
hist(d$zombies_killed)
```

### The Number of zombies killed appears to be a poisson distribution

## Years of Education
```{r}
qqnorm(d$years_of_education, main = "Normal QQ plot Years of education")
qqline(d$years_of_education, col = "gray")
hist(d$years_of_education)
```

### The Number of zombies killed appears to be a poisson distribution


# 2.e. Now use the sample() function to sample ONE subset of 30 zombies (without replacement) from this population and calculate the mean and sample standard deviation for each variable. Also, # estimate the standard error for each variable and construct the 95% confidence interval for each mean. Note that for the variables that are not drawn from the normal distribution, you will need to base your estimate of the CIs on some different distribution!

## First sample the population
```{r}
dsample <- d[sample(1:1000, 30, replace=FALSE), ]
head(dsample)
```

## Calculate mean for each variable
```{r}
mean(dsample$height)
mean(dsample$weight)
mean(dsample$age)
mean(dsample$zombies_killed)
mean(dsample$years_of_education)
```

## Calcualte Standard Deviation for each variable
```{r}
sd(dsample$height)
sd(dsample$weight)
sd(dsample$age)
sd(dsample$zombies_killed)
sd(dsample$years_of_education)
```

### Create function to estimate Standard Error
```{r}
se <- function(x) {
  sd(x)/sqrt(length(x))
}
```

## Estimate Standard error for each variable
```{r}
se(dsample$height)
se(dsample$weight)
se(dsample$age)
se(dsample$zombies_killed)
se(dsample$years_of_education)
```

## Construct 95% confidence intervals for each variable
### HEIGHT
```{r}
qnorm(0.975, mean = mean(dsample$height), sd = sd(dsample$height))
upperheight <- mean(dsample$height) + qnorm(0.975) * se(dsample$height)
lowerheight <- mean(dsample$height) - qnorm(0.975) * se(dsample$height)
ciheight <- c(lowerheight, upperheight)
ciheight



```

### WEIGHT
```{r}
qnorm(0.975, mean = mean(dsample$weight), sd = sd(dsample$weight))
upperweight <- mean(dsample$weight) + qnorm(0.975) * se(dsample$weight)
lowerweight <- mean(dsample$weight) - qnorm(0.975) * se(dsample$weight)
ciweight <- c(lowerweight, upperweight)
ciweight
```

## AGE
```{r}
qnorm(0.975, mean = mean(dsample$age), sd = sd(dsample$age))
upperage <- mean(dsample$age) + qnorm(0.975) * se(dsample$age)
lowerage <- mean(dsample$age) - qnorm(0.975) * se(dsample$age)
ciage <- c(lowerage, upperage)
ciage
```

## ZOMBIES KILLED
```{r}
qpois(0.975, lambda = mean(dsample$zombies_killed))
upperzombies <- mean(dsample$zombies_killed) + qpois(0.975,
      lambda = mean(dsample$zombies_killed)) * se(dsample$zombies_killed)
lowerzombies <- mean(dsample$zombies_killed) - qpois(0.975,
      lambda = mean(dsample$zombies_killed)) * se(dsample$zombies_killed)
cizombies <- c(lowerzombies, upperzombies)
cizombies
```

## YEARS OF EDCUCATION
```{r}
qpois(0.975, lambda = mean(dsample$years_of_education))
upperyears_of_education <- mean(dsample$years_of_education) + qpois(0.975,
      lambda = mean(dsample$years_of_education)) * se(dsample$years_of_education)
loweryears_of_education <- mean(dsample$years_of_education) - qpois(0.975,
      lambda = mean(dsample$years_of_education)) * se(dsample$years_of_education)
cieducation <- c(loweryears_of_education, upperyears_of_education)
cieducation
```

# 2.f. Now draw 99 more random samples of 30 zombies out and calculate the mean for each of the these samples. Together with the first sample you drew out, you now have a set of 100 means for each variable (each based on 30 observations), which constitutes a sampling distribution for each variable. What are the means and standard deviations of the sampling distribution for each variable? How do the standard deviations compare to the standard errors estimated in [2.e.]? What do these sampling distributions look like? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?

## Sample each variable of interest in 99 iterations of 30 samples each from the original dataset
```{r}
k <- 99 
n <- 30  
dsample2 <- NULL  
dsamplemeanheight <- NULL
dsampleSDheight <- NULL
dsampleSEheight <- NULL
dsamplemeanweight <- NULL
dsampleSDweight <- NULL
dsampleSEweight <- NULL
dsamplemeanage <- NULL
dsampleSDage <- NULL
dsampleSEage <- NULL
dsamplemeanzombies <- NULL
dsampleSDzombies <- NULL
dsampleSEzombies <- NULL
dsamplemeaneducation <- NULL
dsampleSDeducation <- NULL
dsampleSEeducation <- NULL
 for (i in 1:99) {
   dsample2 <- sample_n(d, size = n, replace = FALSE)
   dsamplemeanheight[i] <- mean(dsample2$height)
   dsampleSDheight[i] <- sd(dsample2$height)
   dsampleSEheight[i] <- se(dsample2$height)
   dsamplemeanweight[i] <- mean(dsample2$weight)
   dsampleSDweight[i] <- sd(dsample2$weight)
   dsampleSEweight[i] <- se(dsample2$weight)
   dsamplemeanage[i] <- mean(dsample2$age)
   dsampleSDage[i] <- sd(dsample2$age)
   dsampleSEage[i] <- se(dsample2$age)
   dsamplemeanzombies[i] <- mean(dsample2$zombies_killed)
   dsampleSDzombies[i] <- sd(dsample2$zombies_killed)
   dsampleSEzombies[i] <- se(dsample2$zombies_killed)
   dsamplemeaneducation[i] <- mean(dsample2$years_of_education)
   dsampleSDeducation[i] <- sd(dsample2$years_of_education)
   dsampleSEeducation[i] <- se(dsample2$years_of_education)
   }
```

## Create a data frame with means for each of the variables across all of the 99 new samples
```{r}
sampledistmeanheight <- as.data.frame(dsamplemeanheight)
sampledistSDheight <- as.data.frame(dsampleSDheight)
sampledistSEheight <- as.data.frame(dsampleSEheight)
sampledistmeanweight <- as.data.frame(dsamplemeanweight)
sampledistSDweight <- as.data.frame(dsampleSDweight)
sampledistSEweight <- as.data.frame(dsampleSEweight)
sampledistmeanage <- as.data.frame(dsamplemeanage)
sampledistSDage <- as.data.frame(dsampleSDage)
sampledistSEage <- as.data.frame(dsampleSEage)
sampledistmeanzombies <- as.data.frame(dsamplemeanzombies)
sampledistSDzombies <- as.data.frame(dsampleSDzombies)
sampledistSEzombies <- as.data.frame(dsampleSEzombies)
sampledistmeaneducation <- as.data.frame(dsamplemeaneducation)
sampledistSDeducation <- as.data.frame(dsampleSDeducation)
sampledistSEeducation <- as.data.frame(dsampleSEeducation)
```

## Combine the means from each sample into a single data frame
```{r}
sampledist <- cbind(sampledistmeanheight, sampledistSDheight, sampledistSEheight,
      sampledistmeanweight, sampledistSDweight, sampledistSEweight,
      sampledistmeanage, sampledistSDage, sampledistSEage,
      sampledistmeanzombies, sampledistSDzombies, sampledistSEzombies,
      sampledistmeaneducation, sampledistSDeducation, sampledistSEeducation)
```

## Take the means of the original sample vector (from 2.e) and add to the sample distribution. This is the sample distribution of all 100 samples
```{r}

dsamplemeans <- c(mean(dsample$height), sd(dsample$height), se(dsample$height),
                  mean(dsample$weight), sd(dsample$weight), se(dsample$weight),
                  mean(dsample$age), sd(dsample$age), se(dsample$age),
                  mean(dsample$zombies_killed), sd(dsample$zombies_killed), se(dsample$zombies_killed),
                  mean(dsample$years_of_education), sd(dsample$years_of_education), se(dsample$years_of_education))
sampledist <- rbind(sampledist, dsamplemeans)
```

## Determine the mean of means, mean of SD's, and mean of SE's of the sample distribution
### Mean of Means for each variable
```{r}
mean(sampledist$dsamplemeanheight)
mean(sampledist$dsamplemeanweight)
mean(sampledist$dsamplemeanage)
mean(sampledist$dsamplemeanzombies)
mean(sampledist$dsamplemeaneducation)
```

### Mean of SD's for each variable
```{r}
sd(sampledist$dsampleSDheight)
sd(sampledist$dsampleSDweight)
sd(sampledist$dsampleSDage)
sd(sampledist$dsampleSDzombies)
sd(sampledist$dsampleSDeducation)
```

### Mean of SE's for each variable
```{r}
se(sampledist$dsampleSEheight)
se(sampledist$dsampleSEweight)
se(sampledist$dsampleSEage)
se(sampledist$dsampleSEzombies)
se(sampledist$dsampleSEeducation)
```

## How do the sampling distribution SD's compare to the SD's from 2e
```{r}
sd(dsample$height)
mean(sampledist$dsampleSDheight)

sd(dsample$weight)
mean(sampledist$dsampleSDweight)

sd(dsample$age)
mean(sampledist$dsampleSDage)

sd(dsample$zombies_killed)
mean(sampledist$dsampleSDzombies)

sd(dsample$years_of_education)
mean(sampledist$dsampleSDeducation)
```
### ???????????????????????????????

## What do these sampling distributions look like? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?

### Create Histograms and QQ plots of each variable

#### Mean Height
```{r}
hist(sampledist$dsamplemeanheight)
qqnorm(sampledist$dsamplemeanheight, main = "Normal QQ plot Height")
qqline(sampledist$dsamplemeanheight, col = "gray")
```


#### Mean Weight
```{r}
hist(sampledist$dsamplemeanweight)
qqnorm(sampledist$dsamplemeanweight, main = "Normal QQ plot Weight")
qqline(sampledist$dsamplemeanweight, col = "gray")
```



#### Mean Age
```{r}
hist(sampledist$dsamplemeanage)
qqnorm(sampledist$dsamplemeanage, main = "Normal QQ plot Age")
qqline(sampledist$dsamplemeanage, col = "gray")
```


#### Mean Zombies Killed
```{r}
hist(sampledist$dsamplemeanzombies)
qqnorm(sampledist$dsamplemeanzombies, main = "Normal QQ plot Mean # of Zombies Killed")
qqline(sampledist$dsamplemeanzombies, col = "gray")
```


#### Mean Years of Education
```{r}
hist(sampledist$dsamplemeaneducation)
qqnorm(sampledist$dsamplemeaneducation, main = "Normal QQ plot Mean Years of Education")
qqline(sampledist$dsamplemeaneducation, col = "gray")
```


#### SD Height
```{r}
hist(sampledist$dsampleSDheight)
qqnorm(sampledist$dsampleSDheight, main = "Normal QQ plot SD Height")
qqline(sampledist$dsampleSDheight, col = "gray")
```


#### SD Weight
```{r}
hist(sampledist$dsampleSDweight)
qqnorm(sampledist$dsampleSDweight, main = "Normal QQ plot SD Weight")
qqline(sampledist$dsampleSDweight, col = "gray")
```


#### SD Age
```{r}
hist(sampledist$dsampleSDage)
qqnorm(sampledist$dsampleSDage, main = "Normal QQ plot SD Age")
qqline(sampledist$dsampleSDage, col = "gray")
```


#### SD Zombies Killed
```{r}
hist(sampledist$dsampleSDzombies)
qqnorm(sampledist$dsampleSDzombies, main = "Normal QQ plot SD # of Zombies Killed")
qqline(sampledist$dsampleSDzombies, col = "gray")
```


#### SD Years of Education
```{r}
hist(sampledist$dsampleSDeducation)
qqnorm(sampledist$dsampleSDeducation, main = "Normal QQ plot SD Years of Education")
qqline(sampledist$dsampleSDeducation, col = "gray")
```


#### SE Height
```{r}
hist(sampledist$dsampleSEheight)
qqnorm(sampledist$dsampleSEheight, main = "Normal QQ plot SE Height")
qqline(sampledist$dsampleSEheight, col = "gray")
```


#### SE Weight
```{r}
hist(sampledist$dsampleSEweight)
qqnorm(sampledist$dsampleSEweight, main = "Normal QQ plot SE Weight")
qqline(sampledist$dsampleSEweight, col = "gray")
```


SE Age
```{r}
hist(sampledist$dsampleSEage)
qqnorm(sampledist$dsampleSEage, main = "Normal QQ plot SE Age")
qqline(sampledist$dsampleSEage, col = "gray")
```


#### SE Zombies Killed
```{r}
hist(sampledist$dsampleSEzombies)
qqnorm(sampledist$dsampleSEzombies, main = "Normal QQ plot SE # of Zombies Killed")
qqline(sampledist$dsampleSEzombies, col = "gray")
```


#### SE Years of Education
```{r}
hist(sampledist$dsampleSEeducation)
qqnorm(sampledist$dsampleSEeducation, main = "Normal QQ plot SE Years of Education")
qqline(sampledist$dsampleSEeducation, col = "gray")
```
